Statically Typed String Sanitation Inside a Python

### 2014-08-14 sanitation-psp14/sanitation-psp14-short.tex
Remove type annotation from fig 2
Changed check to match tcheck; still not sure how the middle trnaslation of tr-check is supposed to work.

\newcommand{\tcheck}[2]{{\sf check}(#1; #2)}
\renewcommand{\tstr}[1]{{{\sf str}[#1]}}
\newcommand{\preplace}[3]{{\sf preplace}(#1;#2;#3)}
\newcommand{\tconcat}[2]{{\sf concat}(#1;#2)} \newcommand{\concat}[2]{{\sf concat}(#1;#2)} % lots of legacy naming around these parts...

@@

% Meta-theoretic functions
\newcommand{\lsubst}[3]{{\tt lsubst}(#1;#2;#3)} % This used to render lreplace(...) so there're probably mistkes wherever this command was used now.
\newcommand{\lreplace}[3]{{\tt lreplace}(#1; #2; #3)}

% Judgements
\newcommand{\err}{\ {\sf err}}
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

@@

<$e$> ::= 
      $x$ \hfill source terms \alt 
      $\lambda x {:} \sigma . e$ \alt
      $e~(e)$ \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{e}{e}$ \alt
      $\sisubst{r}{e}{e}$ \alt
      $\coerce{r}{e}{x.e}{e}$
\caption{Syntax of $\lambda_{RS}$.}\label{fig:lamrs}
\end{grammar}
\end{figure}

@@

\begin{figure}[t]\label{fig:lambdas}
\small
$\fbox{\inferrule{}{\sctx \vdash S : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Abs]
{\sctx, x : \sigma_1 \vdash e : \sigma_2}
{\sctx \vdash \lambda x.e : \sigma_1 \rightarrow \sigma_2}

\inferrule[S-T-App]
{\sctx \vdash e_1 : \sigma_1 \\ \sctx \vdash e_2 : \sigma_1 \rightarrow \sigma}
{\sctx \vdash e_1 e_2 : \sigma}

@@

\begin{figure}[t]
\small
$\fbox{\inferrule{}{S \sreduces S}}$
$\fbox{\inferrule{}{S \err}}$
\begin{mathpar}
\inferrule[S-E-Abs]                                                             
{ \ }                                                                           
{\lambda x.e \sreduces \lambda x.e}

\inferrule[S-E-App]
{ e_1 \sreduces e_1' \\ e_2 \sreduces e_2' }
{ e_1 e_2 \sreduces e_1' e_2'}

@@

\end{mathpar}
\caption{Big step semantics for $\lambdas$.
Error propagation rules are omitted.}
\label{fig:eeval}
\end{figure}

@@

\addtocounter{figure}{4}
\begin{figure*}[t]\label{fig:trans}

$\fbox{\inferrule{}{\trden{S} = P}}$
\begin{mathpar}

@@

\inferrule[Tr-App]
{\sctx \vdash \trden{e_1} = \iota_1 \\ \sctx \vdash \trden{e_2} = \iota_2}
{\sctx \vdash \trden{e_1e_2} = \iota_1\iota_2}

@@

\inferrule[Tr-Concat]
{ \sctx \vdash \trden{e_1} = \iota_1 \\ \sctx \vdash htrden{e_2} = \iota_2}
{ \sctx \vdash \trden{\rsconcat{e_1}{e_2}} = \tconcat{\iota_1}{\iota_2} }

@@

\inferrule[Tr-SafeCoerce]
{ \ }
{ \sctx \vdash  \trden{ \coerce{r'}{e} } = \trden{e} }

\inferrule[Tr-Check-Ok]
{ \sctx \vdash e : \stringin{r} \\ \lang{r} \subseteq \lang{r'} } 
{ \sctx \vdash  \trden{ \rcheck{r'}{e}{x.e_1}{e_2} } = \trden{[e/x]e_1} }

\inferrule[Tr-Check-NotOk]
{ \sctx \vdash e : \stringin{r} \\ \lang{r} \not \subseteq \lang{r'}}
{ \sctx \vdash  \trden{ \rcheck{r'}{e}{x.e_1}{e_2} } = \trden{\tcheck{r'}{e_2}} }

\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).
The translation is type-directed in the Tr-SafeCoerce case.}
\end{figure*}
\addtocounter{figure}{-5}

@@

\begin{figure}[t]\label{fig:lambdap}
\small
$\fbox{\inferrule{}{\tctx \vdash P : \theta}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \theta$

\begin{mathpar}
\inferrule[P-T-Abs]
{\tctx, x : \iota_1 \vdash e : \iota_2}
{\tctx \vdash \lambda x.e : \iota_1 \rightarrow \iota_2}

\inferrule[P-T-App]
{\tctx \vdash e_1 : \iota_1 \\ \tctx \vdash e_2 : \iota_1 \rightarrow \iota}
{\tctx \vdash e_1 e_2 : \iota}

@@

\inferrule[P-T-Check]
{\tctx \vdash \iota_1 : \regex \\ \tctx \vdash \iota_2 : \str \\ }
{\tctx \vdash \tcheck{\iota_1}{\iota_2} : \str}
\end{mathpar}
\caption{Typing rules for $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:ityping}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{P \treduces P}}$
$\fbox{\inferrule{}{P \err}}$

@@

\inferrule[P-E-App]
{ e_1 \sreduces e_1' \\  e_2 \sreduces e_2' }
{ e_1 e_2 \sreduces e_1' e_2'}

@@

\inferrule[P-E-Check-OK]
{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \strin{s} \\ s \in \lang{r}}
{\tcheck{\iota_1}{\iota_2} \treduces \tstr{s}}

\inferrule[P-E-Check-Err]
{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \tstr{s} \\ s \not \in \lang{r}}
{\tcheck{\iota_1}{\iota_2} \err}
\end{mathpar}
\caption{Big step semantics for of $\lambdap$.
Error propagation rules are omitted.}
\label{fig:ieval}
\end{figure}

\newpage
\ 
\newpage
\ 
\newpage
\

@@

<P> ::= $\lambda x . \iota$ \hfill target terms \alt
  $\iota \iota$ \alt
  $\tstr{s}$ \alt
  $\rx{r}$ \alt
  $\tconcat{\iota}{\iota}$ \alt
  $\preplace{\iota}{\iota}{\iota}$ \alt
  $\tcheck{\iota}{\iota}$ 
  \end{grammar}
\caption{Syntax for the target language, $\lambdap$, containing strings and statically constructed regular expressions.}
\label{fig:lcsSyntax}
\end{figure}

@@

\begin{defn}[$\tt{lsubst}$]
The relation $\lsubst{r}{s_1}{s_2} = s$ produces a string $s$ in which all substrings of $s_1$ matching $r$ are replaced with $s_2$.
\end{defn}

@@

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the definitions of lsubst and lreplace; note that language substitutions over-approximate string substitutions.
\end{proof}

>>>>

\newcommand{\tcheck}[4]{{\sf check}(#1; #2; #3; #4)}
\renewcommand{\tstr}[1]{{{\sf str}[#1]}}
\newcommand{\preplace}[3]{{\sf preplace}(#1;#2;#3)}
\newcommand{\tconcat}[2]{{\sf concat}(#1;#2)} \newcommand{\concat}[2]{{\sf concat}(#1;#2)} % lots of legacy naming around these parts...

@@

% Meta-theoretic functions
\newcommand{\lsubst}[3]{{\tt subst}(#1;#2;#3)} % This used to renderlreplace(...) so there're probably mistkes wherever this command was used now.
\newcommand{\lreplace}[3]{{\tt lreplace}(#1; #2; #3)}

% Judgements
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

@@

<$e$> ::= 
      $x$ \hfill source terms \alt 
      $\lambda x . e$ \alt
      $e~(e)$ \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{e}{e}$ \alt
      $\sisubst{r}{e}{e}$ \alt
      $\rcoerce{r}{e}$ \alt
      $\rcheck{r}{e}{x.e}{e}$
\caption{Syntax of $\lambda_{RS}$.}\label{fig:lamrs}
\end{grammar}
\end{figure}

@@

\begin{figure}[t]\label{fig:lambdas}
\small
$\fbox{\inferrule{}{\sctx \vdash e : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Var]
{ x:\sigma \in \sctx }
{ \sctx \vdash x:\sigma}

\inferrule[S-T-Abs]
{\sctx, x : \sigma_1 \vdash e : \sigma_2}
{\sctx \vdash \lambda x.e : \sigma_1 \rightarrow \sigma_2}

\inferrule[S-T-App]
{\sctx \vdash e_1 : \sigma_2 \rightarrow \sigma \\ \sctx \vdash e_2 : \sigma_2}
{\sctx \vdash e_1(e_2) : \sigma}

@@

\begin{figure}[t]
\small
$\fbox{\inferrule{}{e \sreduces e}}$
\begin{mathpar}
\inferrule[S-E-Abs]                                                             
{ \ }                                                                           
{\lambda x.e \sreduces \lambda x.e}

\inferrule[S-E-App]
{ e_1 \sreduces e_1' \\ e_2 \sreduces e_2' }
{ e_1(e_2) \sreduces e_1'(e_2')}

@@

\end{mathpar}
\caption{Big step semantics for $\lambdas$}.
\label{fig:eeval}
\end{figure}

@@

\begin{figure*}[t]\label{fig:trans}

$\fbox{\inferrule{}{\sctx \vdash \trden{e} = \iota}}$
\begin{mathpar}

@@

\inferrule[Tr-App]
{\sctx \vdash \trden{e_1} = \iota_1 \\ \sctx \vdash \trden{e_2} = \iota_2}
{\sctx \vdash \trden{e_1(e_2)} = \iota_1(\iota_2)}

@@

\inferrule[Tr-Concat]
{ \sctx \vdash \trden{e_1} = \iota_1 \\ \sctx \vdash \trden{e_2} = \iota_2}
{ \sctx \vdash \trden{\rsconcat{e_1}{e_2}} = \tconcat{\iota_1}{\iota_2} }

@@

\inferrule[Tr-SafeCoerce]
{ \sctx \vdash \trden{e} = \iota }
{ \sctx \vdash  \trden{ \coerce{r'}{e} } = \iota }

\inferrule[Tr-Check]
{ \sctx \vdash \trden{e} = \iota_1 \\ \sctx \vdash \trden{e_1} = \iota_1?? \\ \sctx \vdash \trden{e_2} = \iota_2} 
{ \sctx \vdash \trden{ \rcheck{r}{e}{x.e_1}{e_2} } = \tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} }

\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).
The translation is type-directed.}
\end{figure*}

@@

\begin{figure}[t]\label{fig:lambdap}
\small
$\fbox{\inferrule{}{\tctx \vdash \iota : \tau}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \tau$

\begin{mathpar}
\inferrule[P-T-Var]
{ x:\tau \in \tctx }
{ \tctx \vdash x:\tau }

\inferrule[P-T-Abs]
{\tctx, x : \tau_1 \vdash \iota_2 : \tau_2}
{\tctx \vdash \lambda x.\iota_2 : \tau_1 \rightarrow \tau_2}

\inferrule[P-T-App]
{\tctx \vdash \iota_1 : \tau_2 \rightarrow \tau \\ \tctx \vdash \iota_2 : \tau_2}
{\tctx \vdash \iota_1(\iota_2) : \iota}

@@

\inferrule[P-T-Check]
{\tctx \vdash \iota_1 : \regex \\ \tctx, x:\str \vdash \iota_2 : \sigma \\ \tctx \vdash \iota_3 : \sigma}
{\tctx \vdash \tcheck{\iota_1}{x.\iota_2}{\iota_3} : \sigma}
\end{mathpar}
\caption{Typing rules for $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:ityping}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\iota \treduces \iota}}$

@@

\inferrule[P-E-App]
{ \iota_1 \sreduces \iota_1' \\  \iota_2 \sreduces \iota_2' }
{ \iota_1(\iota_2) \sreduces \iota_1'(\iota_2')}

@@

\inferrule[P-E-Check-OK]
{i \treduces \str{s} \\ s \in \lang{r}}
{\tcheck{r}{i}{x.i_1}{i_2} \treduces [\str{s}/x]i_1}

\inferrule[P-E-Check-NotOK]
{i \treduces \str{s} \\ s \not \in \lang{r}}
{\tcheck{r}{i}{x.i_1}{i_2} \treduces i_2}

{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \strin{s} \\ s \in \lang{r}}
{\tcheck{\iota_1}{\iota_2} \treduces \tstr{s}}

\end{mathpar}
\caption{Big step semantics for of $\lambdap$}.
\label{fig:ieval}
\end{figure}

@@

<P> ::= $x$  \hfill target terms \alt
   $\lambda x . \iota$ \alt
   $\iota \iota$ \alt
  $\tstr{s}$ \alt
  $\rx{r}$ \alt
  $\tconcat{\iota}{\iota}$ \alt
  $\preplace{\iota}{\iota}{\iota}$ \alt
  $\tcheck{\rx{r}}{\iota}{\iota}{\iota}$ 
  \end{grammar}
\caption{Syntax for the target language, $\lambdap$, containing strings and statically constructed regular expressions.}
\label{fig:lcsSyntax}
\end{figure}

@@

\begin{defn}[$\tt{subst}$]
The relation $\lsubst{r}{s_1}{s_2} = s$ produces a string $s$ in which all substrings of $s_1$ matching $r$ are replaced with $s_2$.
\end{defn}

@@

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the definitions of subst and lreplace; note that language substitutions over-approximate string substitutions.
\end{proof}
